---
title: "Project1"
format: html
editor: source
---

## Data Processing
### Creating calling all the packages we need but coding so that they are not visable on the quarto document. 
```{r}
#| output: false
# Calling libraries
library(tidyverse)
library(tidycensus)
library(httr)
library(jsonlite)
library(glue)

```

### In this step, we create the URL string, starting with the base URL and incorporating our default variables from get_vals. We set a default geography of state 10 (Delaware) to prevent unintended large API calls.

```{r}
source("src/census_query.R")
census_url
source("test/test_census_query.R")
test_1
test_1()
test_2
test_2()
test_3
test_3()
test_4
test_4()
```

### We're creating a function to access API data, allowing the year to change with a default of 2022, along with a list of valid variables for the PUMS API. We're accessing all valid API variables and subsetting them according to project instructions. We also create a function to check if the variables in the subset are numeric or character.

```{r}
source("src/census_variables.R")
census_url
source("test/test_census_query.R")
test_1
test_1()
test_2
test_2()
test_3
test_3()
test_4
test_4()

```

### We're creating a function that matches valid variables with their corresponding numeric ranges and character items. 

```{r}
library(glue)
source("src/input_validation.R")
census_url
source("test/test_input_validation.R")
test_1
test_1()
test_2
test_2()
test_3
test_3()
test_4
test_4()
test_5
test_5()
test_6
test_6()
```
### We're accessing all valid API variables and subsetting them according to project instructions. We also create a function to check if the variables in the subset are numeric or character. We're creating a function that matches valid variables with their corresponding numeric ranges and character items. We're processing time variables by separating the intervals, updating morning/evening abbreviations for usability, and calculating the midpoint to derive a single time point.

```{r}
source("src/main.R")
census_url
source("test/test_main.R")

```

